Security Metrics for the Android Ecosystem

### 2015-07-28 security_scoring.tex
Minor fixes

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/proportioninsecure}
\caption{Proportion of devices running insecure, maybe secure and secure versions of Android.
Table~\ref{tab:andvulns} lists the \daNumVulnsUsed\ vulnerabilities used, the red vertical lines are caused by their discovery and the most important are annotated.}
\label{fig:proportioninsecure}
\end{figure}
By combining data on critical vulnerabilities in Android and the versions of Android running on devices we can determine which vulnerabilities each device was vulnerable to each day.
We consider a device is \emph{insecure} if it is running a vulnerable version of Android and the device has not received an update which might fix it;
it is \emph{maybe secure} if it is running a vulnerable version but received an update which could have fixed the vulnerability if it contained a backported fix;
and it is \emph{secure} if it is running a secure version.
This allows us to plot Figure~\ref{fig:proportioninsecure}, initially all devices are \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
This means we cannot distinguish between devices which are running a version of Android which is known to be vulnerable from one which may have received a backported fix.
This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
Once \avovuln{zergRush} was discovered in October 2011 then most devices are recorded as \emph{insecure} (red) as they were vulnerable.
The remaining devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded or replaced.
This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received an OS update that did not update to a known-good version of Android but which may still have included a backport of a fix, as the update was made available after the vulnerability was disclosed.
From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are vulnerable.
This means that \daMeanInsecurityPerc\ of devices were classified as \emph{insecure} to at least one critical vulnerability between \daStartDateMonth\ and \daEndDateMonth\ (ignoring devices classed as \emph{maybe secure}).
%This is $1 - f$, the proportion of devices free from critical vulnerabilities, however this is only one component of a more robust score, the FUM score.

@@

We can compute the uncertainty for $f$, $u$ and $m$.
$f$ is computed by taking the total secure device days and dividing it by the total insecure and secure device days.
The total secure device days and total insecure device days are both counting experiments and so their measurement error is their square root~\cite{Taylor1997}.
%Since the numbers involved are large, the uncertainty in $f$ is small.
$u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have square root uncertainties.
$m$ is computed by counting the number of vulnerabilities which affected that entity and which have not yet been fixed on any device we have observed from that entity every day and averaging time.
However, it could be that the entity has released a fix to some devices but we have not yet observed a device with that fix.
So the uncertainty in our measurement is the probability of not having observed a fixed device if a fixed device existed.
We assume that if the fix has been released then at least \daMeanOutstandingProbability\ of devices have the fix.
This represents a trade-off between a proportion so small that the fix has not really been deployed and a reasonable estimate of the error.
This gives an uncertainty of \daMeanOutstandingUncertaintyEquation\ where $n$ is the number of devices contributing to that day's data for each vulnerability outstanding each day.
The Python \texttt{uncertainties} library was used to propagate uncertainties through calculations.
This does not capture systematic errors.
For example, we do not include manufacturer specific vulnerabilities, we expect that performance in fixing manufacturer specific vulnerabilities is strongly correlated with performance fixing vulnerabilities affecting all of Android.

@@

\subsection{Results: Security scores}\label{sec:security_scoring:results}
The scores across the whole of Android are that \daMeanInsecurityPerc\ of devices are exposed to known root exploits.
There are on average \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device.
On average \daUpdatednessPerc\ of devices run the most recent version of Android.
This gives a security score of \daSecurityScore\ out of 10.

However there are a wide variety of scores depending on the source of the device.
There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices\footnote{\url{http://www.howtogeek.com/139391/htg-explains-why-android-geeks-buy-nexus-devices/}}.
Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.

@@

Even within device manufacturers, different models can have very different update behaviours and hence security.
Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \emph{\daSecScoreBestmodel} (\daSecScoreBestmodelScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmodel} (\daSecScoreWorstmodelScore\ out of 10) scoring lowest.
We can then test whether this seems fair by comparing the version data for the highest and lowest scoring models.
Figure~\ref{fig:full_version_comp}c shows the full version distribution for \emph{\daSecScoreWorstmodel}, which we only observe running one version.
Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{HTC Desire HD A9191}, which used to be our worst model and for which we have more historical data; it shows how it received one update at the beginning of 2012, which was deployed fairly rapidly to most devices, but received no further updates.
Figure~\ref{fig:full_version_comp}a shows the same information for \emph{\daSecScoreBestmodel} which received \daSecScoreBestmodelNumFullVersions\ different versions, some of which were only deployed to small numbers of devices, but the distribution for all devices regularly and rapidly transitions from one version to another before ending up on `4.3 JWR66Y'.
Both \emph{Galaxy Nexus} and \emph{HTC Desire HD A9191} device models start off with the full version string of `2.3.3 GRI40' but the \emph{\daSecScoreBestmodel} receives many more updates over the same time period.
Other models from the same manufacturer with similar model names to \emph{HTC Desire HD A9191} do much better such as the \emph{Desire HD}.

@@

We also analysed the \daNumSigOperators\ network operators with a significant presence in our data.
Table~\ref{tab:sec_operator} shows the results \emph{\daSecScoreBestoperator} (\daSecScoreBestoperatorScore\ out of 10) scoring highest and \emph{\daSecScoreWorstoperator} (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
However, the score of a network operator is affected by the manufacturers of the devices which are in use on its network.
This is in turn affected by both the device models a network operator offers to users and upon user's choice of device models.
Hence, having a worse score does not necessarily mean that a network operator is worse, it could be that its users all pick phones from a worse device manufacturer, for example, because they were cheaper.
A network operator could use data from this paper to exclude insecure devices from those offered to consumers.
An added value analysis of network operators, which takes into account the device mix used by users of that network operator, would make it possible to determine whether a network operator is making the situation better or worse by the way it ships updates to users.
However our sample size is too small to do that because while we have significant numbers of devices for each of the \daNumSigModels\ device models (Table~\ref{tab:sec_model}) and for each of the \daNumSigOperators\ network operators (Table~\ref{tab:sec_operator}), we would need a significant number of each model in each network operator.
Since the distribution of devices is unlikely to be uniformly distributed across device models and network operators we estimate that 100\,000 unique devices are required each day for at least a year.
This is not an unobtainable number but it is two orders of magnitude more than is available in \da.
%We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.

@@

\subsection{Sensitivity of scoring metric}
\daTabSpearmanRanks
\daTabChangeInScores
To evaluate whether the ranking of different manufacturers is sensitive to the form of the scoring metric we computed the normalised Spearman's Rank correlation coefficient between the lists ordered using different forms of the scoring metric, this is shown in Table~\ref{tab:spearman_ranks}.
The `equal' metric weights $f$, $u$ and $m$ equally rather than favouring $f$ and makes little difference.
Similarly weighting $u$ or $m$ more highly rather than $f$ makes little difference.
While the $f$, $u$ and $m$ components do have some correlation with the overall FUM score, the rankings produced vary substantially.
Changing the scoring metric also impacts the scores given for each entity Table~\ref{tab:change_in_scores} shows the mean impact on the scores.
This shows that $m$ tends to drag down scores.
\todolater{Give some of the actual orderings as well for the most important ones - no space}
\todolater{Sensitivity of time based vs. global scores - no time}

>>>>

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{figures/proportioninsecure}
\caption{Proportion of devices running insecure, maybe secure and secure versions of Android.
Table~\ref{tab:andvulns} lists the \daNumVulnsUsed\ vulnerabilities used, the red vertical lines are caused by their discovery and the most important are annotated.}
\label{fig:proportioninsecure}
\end{figure}
By combining data on critical vulnerabilities in Android and the versions of Android running on devices we can determine which vulnerabilities each device was vulnerable to each day.
We consider a device is \emph{insecure} if it is running a vulnerable version of Android and the device has not received an update which might fix it;
it is \emph{maybe secure} if it is running a vulnerable version but received an update which could have fixed the vulnerability if it contained a backported fix;
and it is \emph{secure} if it is running a secure version.
This allows us to plot Figure~\ref{fig:proportioninsecure}, initially all devices are \emph{maybe secure} (yellow) since \da\ does not have historical data prior to May 2011.
This means we cannot distinguish between devices which are running a version of Android which is known to be vulnerable from one which may have received a backported fix.
This demonstrates the importance of a longitudinal study: this type of analysis requires years of data.
Once \avovuln{zergRush} was discovered in October 2011 then most devices are recorded as \emph{insecure} (red) as they were vulnerable.
The remaining devices were already running a version of Android which fixed the \avovuln{zergRush} vulnerability and are therefore marked as \emph{secure} (green).
From October 2011 until the discovery of \avovuln[APK_duplicate_file]{APK duplicate file} in February 2013 the graph shows progressive improvement as devices are upgraded or replaced.
This means more and more devices are marked as \emph{secure} because they are now running a secure version of Android, or marked as \emph{maybe secure} because they received an OS update that did not update to a known-good version of Android but which may still have included a backport of a fix, as the update was made available after the vulnerability was disclosed.
From February 2013 onwards regular discovery of critical vulnerabilities ensures that most devices are vulnerable.
Ignoring devices classed as \emph{maybe secure}, we find that on average \daMeanInsecurityPerc\ of devices were classed as \emph{insecure} and \daMeanSecurityPercNominal\ classified as \emph{secure} between \daStartDateMonth\ and \daEndDateMonth.
%This is $1 - f$, the proportion of devices free from critical vulnerabilities, however this is only one component of a more robust score, the FUM score.

@@

We can compute the uncertainty for $f$, $u$ and $m$.
$f$ is computed by taking the total secure device days and dividing it by the total insecure and secure device days.
The total secure device days and total insecure device days are both counting experiments and so their measurement error is their square root~\cite{Taylor1997}.
%Since the numbers involved are large, the uncertainty in $f$ is small.
$u$ is computed by taking the sum of the proportions of devices running the most recent version each day, both the count of devices running the maximum version and total count have square root uncertainties.
$m$ is computed by counting the number of vulnerabilities which affected that entity and which have not yet been fixed on any device we have observed from that entity every day and averaging time.
However, it could be that the entity has released a fix to some devices but we have not yet observed a device with that fix.
So the uncertainty in our measurement is the probability of not having observed a fixed device if a fixed device existed.
We assume that if the fix has been released then at least \daMeanOutstandingProbability\ of devices have the fix.
This represents a trade-off between a proportion so small that the fix has not really been deployed and a reasonable estimate of the error.
This gives an uncertainty of \daMeanOutstandingUncertaintyEquation\ where $n$ is the number of devices contributing to that day's data for each vulnerability outstanding each day.
The Python \texttt{uncertainties} library was used to propagate uncertainties through calculations.
This does not capture systematic errors.
For example, we do not include manufacturer specific vulnerabilities, however we expect that performance in fixing manufacturer specific vulnerabilities is strongly correlated with performance fixing vulnerabilities affecting all of Android.

@@

\subsection{Results: Security scores}\label{sec:security_scoring:results}
On average, between \daStartDateMonth\ and \daEndDateMonth\, we found \daMeanOutstandingVulnerabilities\ outstanding vulnerabilities not fixed on any device and \daUpdatednessPerc\ of devices to run the most recent version of Android.
This gives a security score of \daSecurityScore\ out of 10.

However there are a wide variety of scores depending on the source of the device.
There is anecdotal evidence that Google's Nexus devices are better at getting updates than other Android devices because Google makes the original updates and ships them to its devices.\footnote{\url{http://www.howtogeek.com/139391/htg-explains-why-android-geeks-buy-nexus-devices/}}
Table~\ref{tab:sec_summary} shows that this is the case with Nexus devices getting much better scores than non-Nexus devices.

@@

Even within device manufacturers, different models can have very different update behaviours and hence security.
Table~\ref{tab:sec_model} shows the results for the \daNumSigModels\ device models which have a significant presence by the same metric with \emph{\daSecScoreBestmodel} (\daSecScoreBestmodelScore\ out of 10) scoring highest and \emph{\daSecScoreWorstmodel} (\daSecScoreWorstmodelScore\ out of 10) scoring lowest.
We can then test whether this seems fair by comparing the version data for the highest and lowest scoring models.
Figure~\ref{fig:full_version_comp}c shows the full version distribution for \emph{\daSecScoreWorstmodel}, which we only observe running one version.
Figure~\ref{fig:full_version_comp}b shows the full version distribution for \emph{HTC Desire HD A9191}, which used to be our worst model and for which we have more historical data; it shows it received one update at the beginning of 2012, which was deployed fairly rapidly to most devices, but received no further updates.
Figure~\ref{fig:full_version_comp}a shows the same information for \emph{\daSecScoreBestmodel} which received \daSecScoreBestmodelNumFullVersions\ different versions, some of which were only deployed to small numbers of devices, but the distribution for all devices regularly and rapidly transitions from one version to another before ending up on `4.3 JWR66Y'.
Both \emph{Galaxy Nexus} and \emph{HTC Desire HD A9191} device models start off with the full version string of `2.3.3 GRI40' but the \emph{\daSecScoreBestmodel} receives many more updates over the same time period.
Other models from the same manufacturer with similar model names to \emph{HTC Desire HD A9191} do much better such as the \emph{Desire HD}.

@@

We also analysed the \daNumSigOperators\ network operators with a significant presence in our data.
Table~\ref{tab:sec_operator} shows the results with \emph{\daSecScoreBestoperator} (\daSecScoreBestoperatorScore\ out of 10) scoring highest and \emph{\daSecScoreWorstoperator} (\daSecScoreWorstoperatorScore\ out of 10) scoring lowest.
However, the score of a network operator is affected by the manufacturers of the devices which are in use on its network.
This is in turn affected by both the device models a network operator offers to users and upon user's choice of device models.
Hence, having a worse score does not necessarily mean that a network operator is worse, it could be that its users all pick phones from a worse device manufacturer, for example, because they were cheaper.
A network operator could use data from this paper to exclude insecure devices from those offered to consumers.
An added value analysis of network operators, which takes into account the device mix used by users of that network operator, would make it possible to determine whether a network operator is making the situation better or worse by the way it ships updates to users.
However our sample size is too small to do that because while we have significant numbers of devices for each of the \daNumSigModels\ device models (Table~\ref{tab:sec_model}) and for each of the \daNumSigOperators\ network operators (Table~\ref{tab:sec_operator}), we would need a significant number of each model in each network operator.
Since the distribution of devices is unlikely to be uniformly distributed across device models and network operators we estimate that 100\,000 unique devices are required each day for at least a year.
This is not an unobtainable number but it is two orders of magnitude more than is available in \da.
%We do not attempt to disambiguate users behaviour in whether they install updates from network operators using rolling upgrades.

@@

\subsection{Sensitivity of scoring metric}
\daTabSpearmanRanks
\daTabChangeInScores
To evaluate whether the ranking of different manufacturers is sensitive to the form of the scoring metric we computed the normalised Spearman's Rank correlation coefficient between the lists ordered using different forms of the scoring metric, this is shown in Table~\ref{tab:spearman_ranks}.
In the table, the `equal' metric weights $f$, $u$ and $m$ equally rather than favouring $f$ and makes little difference.
Similarly weighting $u$ or $m$ more highly rather than $f$ makes little difference.
While the $f$, $u$ and $m$ components do have some correlation with the overall FUM score, the rankings produced vary substantially.
Changing the scoring metric also impacts the scores given for each entity Table~\ref{tab:change_in_scores} shows the mean impact on the scores.
This shows that $m$ tends to drag down scores.
\todolater{Give some of the actual orderings as well for the most important ones - no space}
\todolater{Sensitivity of time based vs. global scores - no time}

### 2015-07-28 spsm04-thomas.tex
Drop DOIs for lack of space :-(
Minor fixes

\usepackage{subcaption}
\usepackage{tikz}
\usepackage{siunitx}
\usepackage[disable]{todonotes}%
\let\OldTodo\todo
\renewcommand{\todo}{\OldTodo[inline]}
\newcommand{\todolater}[1]{}%{\todo{#1}}%
\RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
\addbibresource{securityupdates.bib}
\renewcommand{\bibfont}{\small}

@@

In order to improve our understanding, we need to know more about the Android ecosystem as a whole.
It is a complex system with many parties involved in a long multi-stage pipeline~\cite{HTC2013}.
We map and quantify the major players in this space who must collaborate to provide updates~(\S\ref{sec:android_ecosystem}) and determine that inaction~(\S\ref{sec:bottleneck}) by some of the manufacturers (we know of \daNumManufacturers) and network operators (\daNumOperators) means many handsets are vulnerable to critical vulnerabilities.
Understanding this ecosystem is all the more important because device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}.

@@

Security for the Android ecosystem can be deployed at three levels: in an online marketplace, at app installation time on the device, and during app execution. 
Indeed, Google provides its users with security in all these places: through analysis of apps by the Play Store, using the Verify Apps feature on the smartphone at installation time, and by an app sandbox on the smartphone during execution. 
The best place to prevent attacks is by sandboxing the app during execution, since all three attack vectors can be prevented at this level whereas not all users install apps exclusively via the Play Store or enable Verify Apps.
In addition, dynamic code loading and injection attacks cannot be discovered at installation time and can be difficult for a marketplace to detect.
Unfortunately, as we shall see, the security sandbox for Android has known critical vulnerabilities on most devices.
This does not mean these devices are attacked, but that they are susceptible to it.
The likelihood of a successful attack then depends on what apps the user installs and where from, as well as the computer networks the device is connected to and the actions the user takes whilst connected.

@@

We used \da\ to record data on the OS versions of Android installed on devices.
Nappa et al.\ performed a similar analysis using WINE to record data on installed versions of Windows client programs and used cleaned NVD data rather than \avo\ for vulnerability data, they did not assign client applications scores~\cite{Nappa2015}.

We use User-Agent strings from Rwandan internet traffic to examine the version distribution of Android OS versions.
User-Agent strings have been used to investigate the timeliness of web browser updates, with at most 80\% of Firefox users running the most recent version~\cite{Frei2008}.
The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
By way of comparison, Android's update process is manual.
The user is notified an update exists, but further action is required, including downloading the update and rebooting the phone to enable installation.
The phone must have sufficient charge to perform the update and the device itself is rendered inoperable during the update process, two factors that might prevent or delay the update process from taking place.
In our data we are unable to determine why a device is not updated. 
It is possible that many updates arrive at handsets, but are simply not installed.
Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation, which is the major problem at present and \S\ref{sec:security_scoring:results} supports this. Further work is required to tease these numbers apart.
Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
Google is deploying the same silent update technique through Google Play Services,\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this bypasses the device manufacturer and network operator.

@@

\section*{Dataset}
Much of the raw and processed data and source code is available, excluding that which might identify individuals~\cite{Thomas2015e}.
Data from \da\ and \avo\ used in this paper is already available.
The Rwanda and FTSE data cannot be made available due to the terms they were obtained under.

>>>>

\usepackage{subcaption}
\usepackage{tikz}
\usepackage{siunitx}
\usepackage[disable]{todonotes}%
\let\OldTodo\todo
\renewcommand{\todo}{\OldTodo[inline]}
\newcommand{\todolater}[1]{}%{\todo{#1}}%
\RequirePackage[date=terse, isbn=true, doi=false, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
\addbibresource{securityupdates.bib}
\renewcommand{\bibfont}{\small}

@@

In order to improve our understanding, we need to know more about the Android ecosystem as a whole.
It is a complex system with many parties involved in a long multi-stage pipeline~\cite{HTC2013}.
We map and quantify the major players in this space who must collaborate to provide updates~(\S\ref{sec:android_ecosystem}) and determine that inaction~(\S\ref{sec:bottleneck}) by some of the manufacturers and network operators means many handsets are vulnerable to critical vulnerabilities.
Understanding this ecosystem is all the more important because device manufacturers have introduced additional vulnerabilities in the past~\cite{Grace2012}.

@@

Security for the Android ecosystem can be deployed at three levels: in an online marketplace, at app installation time on the device, and during app execution. 
Google provides its users with security in all these places: through analysis of apps by the Play Store, using the Verify Apps feature on the smartphone at installation time, and by an app sandbox on the smartphone during execution. 
The best place to prevent attacks is by sandboxing the app during execution, since all three attack vectors can be prevented at this level, whereas not all users install apps exclusively via the Play Store or enable Verify Apps.
In addition, dynamic code loading and injection attacks cannot be discovered at installation time and can be difficult for a marketplace to detect.
Unfortunately, as we shall see, the security sandbox for Android has known critical vulnerabilities on most devices.
This does not mean these devices are attacked, but that they are vulnerable.
The likelihood of a successful attack then depends on what apps the user installs and where from, as well as the computer networks the device is connected to and the actions the user takes whilst connected.

@@

We used \da\ to record data on the OS versions of Android installed on devices.
Nappa et al.\ performed a similar analysis using WINE to record data on installed versions of Windows client programs and used cleaned NVD data rather than \avo\ for vulnerability data.
They did not assign client applications scores~\cite{Nappa2015}.

We use User-Agent strings from Rwandan internet traffic to examine the version distribution of Android OS versions.
User-Agent strings have been used to investigate the timeliness of web browser updates, with at most 80\% of Firefox users running the most recent version~\cite{Frei2008}.
The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
By way of comparison, Android's update process is manual.
The user is notified an update exists, but further action is required, including downloading the update and rebooting the phone to enable installation.
The phone must have sufficient charge to perform the update and the device itself is rendered inoperable during the update process, two factors that might prevent or delay the update process from taking place.
In our data we are unable to determine why a device is not updated. 
It is possible that many updates arrive at handsets, but are simply not installed.
Anecdotal evidence suggests that it is the lack of updates rather the lack of installation, which is the major problem at present.
Our analysis in \S\ref{sec:bottleneck} supports this view.
Nevertheless it is problematic that an operating system update requires a reboot.
Chrome installs the new version side by side with the old one and switches the next time it is restarted.
The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
Google is deploying the same silent update technique through Google Play Services,\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this bypasses the device manufacturer and network operator.

@@

\section*{Dataset}
Much of the raw and processed data and source code is available, excluding that which might identify individuals~\cite{Thomas2015e}.
Data from \da\ and \avo\ used in this paper is already available.
The Rwanda and FTSE data cannot be made available.
